---
title: "emileygz_OriginalHomeworkCode_04"
author: "Emiley Garcia-Zych"
date: "`r Sys.Date()`"
output: html_document
---

```{r}
library(ggplot2) #load the necessary packages
library(curl)
```

# Homework 4: What's Your Malfunction?

## Emiley Garcia-Zych

1)  Write a simple R function, Z.prop.test(), that can perform one- or two- sample Z tests for proportion data, using the following guidelines.

```{r}
Z.prop.test <- function(p1, n1, p2 = NULL, n2 = NULL, p0, alternative = "two.sided", conf.level = 0.95) { # defines a new fucntion "Z.prop.test()" that takes the arguments "p1" and "p2", sample sizes "n1" and "n2", null hypothesis proportion "p0", the type of test, and confidence level of 95%. If p2 and n2 are not provided, the test will be able to run a one-sample test. 
  
  one_sample_test <- is.null(p2) || is.null(n2) # Checks if either p2 or n2 is NULL to determine one-sample or two-sample test
  
  # Calculate the estimated proportion in a two-sample case
  if (one_sample_test) {
    p1_hat <- p1
  } 
  else {
    p1_hat <- p1
    p2_hat <- p2
  }
  
  # Check rules of thumb
  if (n1 * p1_hat > 5 && n1 * (1 - p1_hat) > 5 && (one_sample_test || (n2 * p2_hat > 5 && n2 * (1 - p2_hat) > 5))) { ##makes sure both samples are large enough to run the tests. 
    standard_error <- sqrt(p1_hat * (1 - p1_hat) / n1) # Calculate the standard error
    
    if (!one_sample_test) { #checks to see if it is a two-sample test. If it is a one-sample test, it continues to the else...
      standard_error2 <- sqrt(p2_hat * (1 - p2_hat) / n2)
      pooled_standard_error <- sqrt((standard_error^2 / n1) + (standard_error2^2 / n2))
      Z_score <- (p1_hat - p2_hat) / pooled_standard_error ##calculates Z statistic 
    } 
    else {
      Z_score <- (p1_hat - p0) / standard_error
    }
    
    if (alternative == "greater") {
      P <- 1 - pnorm(Z)
    } else if (alternative == "less") {
      P <- pnorm(Z)
    } else {
      P <- 2 * (1 - pnorm(abs(Z)))
    }
    
    # Calculate the confidence interval
    alpha <- 1 - conf.level
    if (one_sample_test) {
      margin_of_error <- qnorm(1 - alpha / 2) * standard_error
      CI <- c(p1_hat - margin_of_error, p1_hat + margin_of_error)
    } else {
      margin_of_error <- qnorm(1 - alpha / 2) * pooled_standard_error
      CI <- c(p1_hat - p2_hat - margin_of_error, p1_hat - p2_hat + margin_of_error)
    } 
    
    result <- list(Z = Z, P = P, CI = CI)
    return(result)
  } else { ##returns a warning and lets the user know there was some issue with the data. 
    warning("The rules of thumb (n*p > 5 and n*(1-p) > 5) are violated. The results may not be valid.")
    return(NULL)
  }
}
  
```

2)  The dataset from Kamilar and Cooper has in it a large number of variables related to life history and body size. For this exercise, the end aim is to fit a simple linear regression model to predict longevity (MaxLongevity_m) measured in months from species' brain size (Brain_Size_Species_Mean) measured in grams. Do the following for both longevity\~brain size and log(longevity)\~log(brain size):

```{r}
k_and_c <- curl("https://raw.githubusercontent.com/fuzzyatelin/fuzzyatelin.github.io/master/AN588_Fall23/KamilarAndCooperData.csv")
d <- read.csv(k_and_c, header = TRUE, stringsAsFactors = FALSE)
d
```

```{r}
is.atomic(d) ## I was having trouble and kept running into the atomic value error. 
```

```{r}
getElement(d, 'MaxLongevity_m')
```

a)  Fit the regression model and, using {ggplot2}, produce a scatterplot with the fitted line superimposed upon the data. Append the the fitted model equation to your plot (HINT: use the function geom_text()).

```{r}
library(ggplot2)

# Fit longevity~brain size model
model1 <- lm(d$MaxLongevity_m ~ d$Brain_Size_Species_Mean, data = d)

plot1 <- ggplot(d, aes((d$Brain_Size_Species_Mean), (d$MaxLongevity_m), label = rownames(d))) + geom_smooth(method = "lm", formula = y ~ x, color = "pink") 
plot1 + geom_text(check_overlap = TRUE)

```

```{r}
# Fit log(longevity)~log(brain size) model
model2 <- lm(log(d$MaxLongevity_m) ~ log(d$Brain_Size_Species_Mean), data = d)

plot2 <- ggplot(d, aes(log(Brain_Size_Species_Mean), log(MaxLongevity_m), label = rownames(d)))+ geom_smooth(method = "lm", formula = y ~ x, color = "red") 

plot2 + geom_text(check_overlap = TRUE)

```

b) Identify and interpret the point estimate of the slope (β1), as well as the outcome of the test associated with the hypotheses H0: β1 = 0; HA: β1 ≠ 0. Also, find a 90 percent CI for the slope (β1) parameter.

```{r}
# Assuming you have fitted a linear regression model, e.g., model1

# Point estimate of the slope (β1)
slope_estimate <- coef(model1)["Brain_Size_Species_Mean"]

# Conduct the hypothesis test and find the p-value
p_value <- summary(model1)$coefficient["Brain_Size_Species_Mean"]

# 90% confidence interval for the slope (β1)
conf_interval <- confint(model1, level = 0.90)["Brain_Size_Species_Mean"]

# Interpretation
##cat("Point estimate of the slope (β1):", slope_estimate, "\n")
##cat("Hypothesis test (β1 ≠ 0):\n")
##cat("   - p-value:", p_value, "\n")
##if (p_value < 0.05) {
##  cat("   - Interpretation: We reject the null hypothesis (β1 = 0) at the 0.05 significance level.\n")
##} else {
##  cat("   - Interpretation: We do not have sufficient evidence to reject the null hypothesis (β1 = 0) at the 0.05 significance level.\n")
##}
##cat("90% Confidence Interval for the slope (β1): [", conf_interval[1], ",", conf_interval[2], "]\n")

```


```{r}
summary(model1)
summary(model2)
```

```{r}
confint(model1, level = 0.9)
confint(model2, level = 0.9)
```

```{r}
# Predict longevity for species with brain weight of 800 gm using the first model
new_data <- data.frame(Brain_Size_Species_Mean = 800)
predicted_longevity <- predict(model1, newdata = new_data)

# 90% prediction interval
pi <- predict(model1, newdata = new_data, interval = "prediction", level = 0.9)

print("Predicted Longevity:")
print(predicted_longevity)
print("90% Prediction Interval:")
print(pi)
```
