---
title: "emileygz_OriginalHomeworkCode_04"
author: "Emiley Garcia-Zych"
date: "`r Sys.Date()`"
output: html_document
---
```{r}
install.packages("ggplot2")
```


1) Write a simple R function, Z.prop.test(), that can perform one- or two- sample Z tests for proportion data, using the following guidelines. 

```{r}
Z.prop.test <- function(p1, n1, p2 = NULL, n2 = NULL, p0, alternative = "two.sided", conf.level = 0.95) {
  # Check if either p2 or n2 is NULL to determine one-sample or two-sample test
  one_sample_test <- is.null(p2) || is.null(n2)
  
  # Calculate the estimated proportion in the two-sample case
  if (one_sample_test) {
    p1_hat <- p1
  } else {
    p1_hat <- p1
    p2_hat <- p2
  }
  
  # Check rules of thumb
  if (n1 * p1_hat > 5 && n1 * (1 - p1_hat) > 5 && (one_sample_test || (n2 * p2_hat > 5 && n2 * (1 - p2_hat) > 5))) {
    # Calculate the standard error
    se <- sqrt(p1_hat * (1 - p1_hat) / n1)
    
    if (!one_sample_test) {
      se2 <- sqrt(p2_hat * (1 - p2_hat) / n2)
      pooled_se <- sqrt((se^2 / n1) + (se2^2 / n2))
      Z <- (p1_hat - p2_hat) / pooled_se
    } else {
      Z <- (p1_hat - p0) / se
    }
    
    if (alternative == "greater") {
      P <- 1 - pnorm(Z)
    } else if (alternative == "less") {
      P <- pnorm(Z)
    } else {
      P <- 2 * (1 - pnorm(abs(Z)))
    }
    
    # Calculate the confidence interval
    alpha <- 1 - conf.level
    if (one_sample_test) {
      margin_of_error <- qnorm(1 - alpha / 2) * se
      CI <- c(p1_hat - margin_of_error, p1_hat + margin_of_error)
    } else {
      margin_of_error <- qnorm(1 - alpha / 2) * pooled_se
      CI <- c(p1_hat - p2_hat - margin_of_error, p1_hat - p2_hat + margin_of_error)
    }
    
    result <- list(Z = Z, P = P, CI = CI)
    return(result)
  } else {
    warning("The rules of thumb (n*p > 5 and n*(1-p) > 5) are violated. The results may not be valid.")
    return(NULL)
  }
}

  
```
2) The dataset from Kamilar and Cooper has in it a large number of variables related to life history and body size. For this exercise, the end aim is to fit a simple linear regression model to predict longevity (MaxLongevity_m) measured in months from species’ brain size (Brain_Size_Species_Mean) measured in grams. Do the following for both longevity~brain size and log(longevity)~log(brain size):

```{r}
library(curl)
data <- curl("https://raw.githubusercontent.com/fuzzyatelin/fuzzyatelin.github.io/master/AN588_Fall23/KamilarAndCooperData.csv")
d <- read.csv(data, header = TRUE, sep = ",", stringsAsFactors = FALSE)
d
```



```{r}
is.atomic(d$MaxLongevity_m)
```


```{r}
# Fit longevity~brain size model
model1 <- lm(d$MaxLongevity_m ~ d$Brain_Size_Species_Mean, data = d)
```


```{r}
# Fit log(longevity)~log(brain size) model
model2 <- lm(log(d$MaxLongevity_m) ~ log(d$Brain_Size_Species_Mean), data = d)
```


```{r}
library(ggplot2)
# Create a scatterplot
scatterplot1 <- ggplot(d, aes(x = Brain_Size_Species_Mean, y = MaxLongevity_m)) +
  geom_point() +  # Add points
  geom_smooth(method = "lm", formula = y ~ x, color = "blue", se = FALSE) +  # Add the fitted line
  geom_text(aes(label = paste("y = ", round(coef(model1)[1], 2), " + ", round(coef(model1)[2], 2), "x")), 
            x = min(d$Brain_Size_Species_Mean), y = max(d$MaxLongevity_m), 
            hjust = 0, vjust = 1, color = "blue") +  # Add the equation to the plot
  labs(title = "Scatterplot: Longevity ~ Brain Size",
       x = "Brain Size (grams)",
       y = "Longevity (months)")

print(scatterplot1)
```
```{r}
# Create scatterplot for log(longevity)~log(brain size)
plot2 <- ggplot(d, aes(x = log(Brain_Size_Species_Mean), y = log(MaxLongevity_m))) +
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ x, color = "green", se = FALSE) +
  geom_text(aes(label = paste("log(y) = ", round(coef(model2)[1], 2), " + ", round(coef(model2)[2], 2), "log(x)")), 
            x = min(log(d$Brain_Size_Species_Mean)), y = max(log(d$MaxLongevity_m)), 
            hjust = 0, vjust = 1, color = "green") +
  labs(title = "log(Longevity) ~ log(Brain Size)", x = "log(Brain Size)", y = "log(Longevity)")
plot2
```

```{r}
summary(model1)
summary(model2)
```
```{r}
confint(model1, level = 0.9)
confint(model2, level = 0.9)
```
```{r}
# Predict longevity for species with brain weight of 800 gm using the first model
new_data <- data.frame(Brain_Size_Species_Mean = 800)
predicted_longevity <- predict(model1, newdata = new_data)

# 90% prediction interval
pi <- predict(model1, newdata = new_data, interval = "prediction", level = 0.9)

print("Predicted Longevity:")
print(predicted_longevity)
print("90% Prediction Interval:")
print(pi)
```


